---
title: "Exercise 1: Central Limit Theorem"
output: html_document
---

The central limit theorem states that the sum or average of random samples from any probability distribution will approximate a normal distribution, with a sufficient number of samples. This exercise should help you getting an intuition what a "sufficient number" of samples is and what this implies for the assumption of the normal distribution in experimental settings.

Your task is to:

1. choose any - ideally asymmetric or totally random (e.g. using the `sample` function) - probability distribution
2. compute the average of 3, 5, 10, and 20 samples from the distribution for 200 times. 
3. Plot these averages in a histogram and try overlaying the density of the normal distribution with the mean and standard deviation of the plotted values


## Solution

We first create a vector of probabilities that sums to one. I generated random values from a normal distribution and used the softmax function that normalizes a vector of n values into a probability vector of n+1 values. Then I sampled a given number of values from the population of 1:(n+1) integer values

```{r}
my_probs <- bmm::softmax(rnorm(4))
my_samples <- sample(1:5, size = 3, replace = T, prob = my_probs)
```

Then we can calculate the average for these sampled values

```{r}
average = mean(my_samples)
```


Now, we have to repeat this for 200 times

```{r}
n_rep <- 200

sim_results <- data.frame(
  sim_run = 1:n_rep,
  avg_sample3 = numeric(n_rep),
  avg_sample5 = numeric(n_rep),
  avg_sample10 = numeric(n_rep),
  avg_sample20 = numeric(n_rep)
)

for (i in 1:200) {
  my_samples <- sample(1:5, size = 20, replace = T, prob = my_probs)
  
  sim_results$avg_sample3[i] <- mean(my_samples[1:3])
  sim_results$avg_sample5[i] <- mean(my_samples[1:5])
  sim_results$avg_sample10[i] <- mean(my_samples[1:10])
  sim_results$avg_sample20[i] <- mean(my_samples[1:20])
}
```

Let's reshape the simulated data to make plotting a little easier, and the plot histograms.

```{r}
library(tidyverse)

df_results_plot <- sim_results %>% 
  pivot_longer(cols = starts_with("avg_")) %>% 
  mutate(sample_size = str_split_i(name,"_",2))

average_stat_sim <- df_results_plot %>% 
  summarise(
    mean = mean(value),
    sd = sd(value),
    .by = "sample_size"
  )

ggplot(data = df_results_plot,
       aes(x = value)) +
  facet_wrap(vars(sample_size)) +
  geom_histogram(aes(y = after_stat(density)))
```

