---
title: "Analyzing data on the level of cognitive processes"
output:
  pdf_document: default
  html_document: default
---

# Exercise 3 - Basic M3 Model

In this exercise we will implement the basic M3 model to a data sets from Oberauer (2019) "Working Memory Capacity Limits Memory for Bindings" in which participants completed a cued recall task. In two experiments, participant were asked to memorize list. In different conditions, various numbers of words (set size 2, 4, 6, 8) were displayed within a list. At test, participants were cued with a list position and asked to select the word from a selection of words. This selection could contain apart from the correct word, other presented list words or completely new words.

Here is an overview what the individual variable in the data set refer to:

-   id: participant ID
-   setsize: setsize if words within a list (2, 4, 6 or 8) as a factor
-   setsize_lin: setsize if words within a list (2, 4, 6 or 8) as a linear variable
-   n_IIP: number of correct responses (IIP)
-   n_IOP: number of errors when a word from the list was reported, but at the incorrect position (IOP)
-   npl: number of errors when a word *not* from the list was reported (NPL)
-   dist: number of distractor words that were reported instead of a memory list word (only given in the *"new distractors"* condition)
-   n_IIP: total number of correct responses in each trial
-   n_IOP: total number of IOP responses in each trial
-   n_NPL: total number of NPL in each trial

## Task 1: Basic M3 model

We are now interested in investigating the influence of the processing task on memorizing the words and how different cognitive processes were impacted, namely binding memory and item memory. Specifically, how well did participants remember the memory list words when the same word was displayed in the processing task (*"old same"*) in comparison when it was not (*"old reordered"*)?

To this end, we will implement the basic M3 model and will focus on the conditions *"old reordered"* and *"old same"*. How are binding memory and item memory affected by the different distractor conditions? Does this match your expectations?

### Load libraries

```{r}
rm(list = ls())

library("here")
library("bmm")
library("dplyr")
library("tidyr")
library("brms")
library("ggplot2")
library("tidybayes")
library("readr")
```

### Read in data

```{r}
mydata_exp1 <- read.table(here("data", "exc3_Oberauer_2019_E1.txt"))

mydata_exp1 <- mydata_exp1 %>% 
  rename(IIP = correct,
         IOP = intrusion,
         NPL = new,
         n_IIP = rsizeCorrect,
         n_IOP = rsizeList,
         n_NPL = rsizeNPL,
         setsize_lin = ss_lin) %>% 
  mutate(setsize = as.factor(setsize)) %>% 
  select(-ss_fac)

mydata_exp2 <- read.table(here("data", "exc3_Oberauer_2019_E2.txt"))

mydata_exp2 <- mydata_exp2 %>% 
  rename(IIP = correct,
         IOP = intrusion,
         NPL = new,
         n_IIP = rsizeCorrect,
         n_IOP = rsizeList,
         n_NPL = rsizeNPL) %>% 
  mutate(setsize = as.factor(setsize))

mydata_exp1 %>%  
  group_by(id, setsize, rsizeCorrect, rsizeList, rsizeNPL) %>% 
  summarise(n = n())
```

### Implement basic M3 model

```{r}

n_cores <- 4
n_iters <- 2000
n_warmup <- 1000
n_chains <- 4

m3_model_ss <- m3(resp_cats = c("IIP", "IOP", "NPL"),
               num_options = c("n_IIP", "n_IOP", "n_NPL"),
               choice_rule = "softmax",
               version = "ss")

m3_formula_ss <- bmf(
  IIP ~ b + a + c,
  IOP ~ b + a,
  NPL ~ b,
  c ~ 1 + setsize + (1 + setsize | id),
  a ~ 1 + setsize + (1 + setsize | id)
)

default_prior(m3_formula_ss, data = mydata_exp1, model = m3_model_ss)

m3_fit_ss <- bmm(
  formula = m3_formula_ss,
  data = mydata_exp1,
  model = m3_model_ss,
  core = n_cores,
  chain = n_chains,
  iter = n_warmup + n_iters,
  warmup = n_warmup,
  file = here("models", "model_m3_ss"),
  file_refit = "never"
)

summary(m3_fit_ss)
plot(m3_fit_ss)

posterior <- as_draws_df(m3_fit_ss) %>% 
  select(starts_with("b_")) %>% 
  pivot_longer(cols = starts_with("b_"), names_to = "effect", values_to = "estimate") %>% 
  filter(effect != "b_b_Intercept")

ggplot(posterior, aes(x = effect, y = estimate)) +
  stat_slab()

conditional_effects(m3_fit_ss, categorical = T, effects = "setsize")

m3_model_ss2 <- m3(resp_cats = c("IIP", "IOP", "NPL"),
               num_options = c("n_IIP", "n_IOP", "n_NPL"),
               choice_rule = "simple",
               version = "ss")

m3_formula_ss2 <- bmf(
  IIP ~ b + a + c,
  IOP ~ b + a,
  NPL ~ b,
  c ~ 1 + ss_lin + (1 + ss_lin | id),
  a ~ 1 + ss_lin + (1 + ss_lin | id)
)

default_prior(m3_formula_ss2, data = mydata_exp1, model = m3_model_ss)

m3_fit_ss2 <- bmm(
  formula = m3_formula_ss,
  data = mydata_exp1,
  model = m3_model_ss2,
  core = n_cores,
  chain = n_chains,
  iter = n_warmup + n_iters,
  warmup = n_warmup,
  file = here("models", "model_m3_ss2"),
  file_refit = "always"
)

conditional_effects(m3_fit_ss2, categorical = T, effects = "setsize")


```

## Task 2: Extended M3 model

```{r}

n_cores <- 4
n_iters <- 2000
n_warmup <- 1000
n_chains <- 4

m3_model_ss2 <- m3(resp_cats = c("IIP", "IOP", "NPL"),
               num_options = c("n_IIP", "n_IOP", "n_NPL"),
               choice_rule = "softmax",
               version = "ss")

m3_formula_ss2 <- bmf(
  IIP ~ b + a + c,
  IOP ~ b + a,
  NPL ~ b,
  c ~ 1 + setsize + (1 + setsize | id),
  a ~ 1 + setsize + (1 + setsize | id)
)

default_prior(m3_formula_ss2, data = mydata_exp2, model = m3_model_ss)

m3_fit_ss2 <- bmm(
  formula = m3_formula_ss2,
  data = mydata_exp2,
  model = m3_model_ss2,
  core = n_cores,
  chain = n_chains,
  iter = n_warmup + n_iters,
  warmup = n_warmup,
  file = here("models", "model_m3_ss2"),
  file_refit = "never"
)

summary(m3_fit_ss2)
plot(m3_fit_ss2)

posterior <- as_draws_df(m3_fit_ss2) %>% 
  select(starts_with("b_")) %>% 
  pivot_longer(cols = starts_with("b_"), names_to = "effect", values_to = "estimate") %>% 
  filter(effect != "b_b_Intercept")

ggplot(posterior, aes(x = effect, y = estimate)) +
  stat_slab()

conditional_effects(m3_fit_ss2, categorical = T, effects = "setsize")


```