---
title: "Analyzing data on the level of cognitive processes - ESCoP 2025"
author: "Gidon T. Frischkorn, Ruhi Bhanap, and Isabel Courage"
output:
  pdf_document: default
  html_document: default
---
# Exercise 2: Hierarchical Models & Signal Detection Theory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r packages}
# empty environment
rm(list = ls())

# load libraries
library(brms)
library(tidyverse)
library(here)
```
## A - Fit Bayesian GLMs to the aggregated data

1. Load the aggregated data of Williams et al. (2022) `exercise2_Williams2022_exp2_aggregated.txt`. 

This data set contains aggregated performance indices from a Change Detection Task with Set Size 8 and a between subject manipulation `isAdaptive`. The manipulation entailed a prompt pushing participants to respond equally often with `same` and `different` to achieve a less biased response pattern.

```{r load_data}
data_Williams2022_agg <- read.table(here("data","exercise2_Williams2022_exp2_aggregated.txt"))
head(data_Williams2022_agg)
```

2. Fit Bayesian GLMs with either proportion correct, Cowan's K and dprime as the dependent variable

Use the `brm` function to specify that the DV should vary as a function of the `isAdaptive` between subject manipulation. Set appropriate priors for the `b` coefficients for the respective models. Use the `default_prior` function to get information which priors are set as a default by `brms`.
Set the `sample_prior` option to `TRUE` when fitting the model.

```{r GLM_propCorrect}
# default_prior(prop_correct ~ isAdaptive, data = data_Williams2022_agg)

my_priors_pc <- prior(normal(0,0.25), class = b)

fit_pc <- brm(
  formula = bf(prop_correct ~ 1 + isAdaptive, sigma ~ 1 + isAdaptive),
  data = data_Williams2022_agg,
  family = gaussian(),
  cores = 4,
  prior = my_priors_pc,
  sample_prior = TRUE,
  backend = "cmdstanr",
  refresh = 0
)
```

```{r GLM_CowansK}
default_prior(K ~ isAdaptive, data = data_Williams2022_agg)

my_priors_CowansK <- prior(normal(0,1), class = b)

fit_CowansK <- brm(
  formula = bf(K ~ 1 + isAdaptive, sigma ~ 1 + isAdaptive),
  data = data_Williams2022_agg,
  family = gaussian(),
  cores = 4,
  prior = my_priors_CowansK,
  sample_prior = TRUE,
  backend = "cmdstanr",
  refresh = 0
)
```

```{r GLM_dprime}
default_prior(dprime ~ isAdaptive, data = data_Williams2022_agg)

my_priors_dprime <- prior(normal(0,1), class = b)

fit_dprime <- brm(
  formula = bf(dprime ~ 1 + isAdaptive, sigma ~ 1 + isAdaptive),
  data = data_Williams2022_agg,
  family = gaussian(),
  cores = 4,
  prior = my_priors_dprime,
  sample_prior = TRUE,
  backend = "cmdstanr",
  refresh = 0
)
```

3. Quantify the strength of evidence for or against an effect of the manipulation

The `hypothesis` function allows to specify to-be-tested hypothesis for any GLM fitted with `brms`. For this you need to pass the `brmfit` object and a vector of character strings specifying the hypothesis you want to test.

```{r hypothesis_pc}
hypothesis(fit_pc,
           "isAdaptiveyes = 0")
```

```{r hypothesis_CowansK}
hypothesis(fit_CowansK,
           "isAdaptiveyes = 0")
```

```{r hypothesis_dprime}
hypothesis(fit_dprime,
           "isAdaptiveyes = 0")
```

## B - Fit Bayesian-hierarchical SDT model to the data

1. Load the non-aggregated data of Williams et al. (2022): `exercise2_Williams2022_exp2.txt`

This data contains more detailed data on the number of hits and false alarms for trials with targets and foils as test stimuli.

```{r load_data2}
data_Williams2022 <- read.table(here("data","exercise2_Williams2022_exp2.txt"))
head(data_Williams2022)
```

2. Specify the non-linear formula to fit an SDT model with this data

Use the `brmsformula` function to specify the non-linear formula implementing an SDT model in a GLM. Predict both `dprime` and the `criterion` by the `isAdaptive` manipulation.

```{r sdt_formula}
sdt_formula <- bf(
  sum_saysame | trials(n_trials) ~ -crit + was_same_trial * dprime,
  dprime ~ 1 + isAdaptive + (1 | id),
  crit ~ 1 + isAdaptive + (1 | id),
  nl = TRUE
)
```

3. Fit the SDT model to the data

Pass the `brmsformula` to the `brm` function and select to fit a binomial model with a probit link function. Again consider specifying adequate priors for the parameters. And use the `hypothesis` function to test on which parameters of the SDT model the `isAdaptive` manipulation has an effect.

```{r priors_sdt}
default_prior(sdt_formula, data = data_Williams2022, family = binomial(link = "probit"))

my_priors_sdt <- prior(normal(0,1), class = b, nlpar = crit) +
  prior(normal(0,1), class = b, nlpar = dprime)
```

```{r fit_sdt}
fit_sdt <- brm(
  formula = sdt_formula,
  data = data_Williams2022,
  family = binomial(link = "probit"),
  prior = my_priors_sdt,
  sample_prior = T,
  cores = 4,
  refresh = 0,
  backend = "cmdstanr"
)

# plot model predicted effects
conditional_effects(fit_sdt, nlpar = "dprime", effects = "isAdaptive")
conditional_effects(fit_sdt, nlpar = "crit", effects = "isAdaptive")

# test if manipulation had effects on SDT parameters
hypothesis(fit_sdt,
           c("dprime_isAdaptiveyes = 0",
             "crit_isAdaptiveyes = 0"))
```



