---
title: "Analyzing data on the level of cognitive processes - ESCoP 2025"
author: "Gidon T. Frischkorn, Ruhi Bhanap, and Isabel Courage"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Solutions: Exercise 2: Hierarchical Models & Signal Detection Theory

```{r packages}
# empty environment
rm(list = ls())

# load libraries
library(brms)
library(tidyverse)
library(here)
```

## A - Fit Bayesian GLMs to the aggregated data

### 1)  Load the **aggregated data** of experiment 2 of Williams et al. (2022).

This data set (filename: `exercise2_Williams2022_exp2_aggregated.txt`) contains aggregated performance indices from a Change Detection Task with Set Size 8 and a between subject manipulation `isAdaptive`. The manipulation entailed a prompt pushing participants to respond equally often with `same` and `different` to achieve a less biased response pattern.

```{r load_data}
data_Williams2022_agg <- read.table(here("data","exercise2_Williams2022_exp2_aggregated.txt"))
head(data_Williams2022_agg)
```

### 2)  Fit Bayesian GLMs with either proportion correct, Cowan's K or dprime as the dependent variable.

Use the `brm` function to specify that the dependent variable should vary as a function of the `isAdaptive` between-subject manipulation. Set appropriate priors for the `b` coefficients for the respective models. Use the `default_prior` function to get information which priors are set as a default by `brms`. Set the `sample_prior` option to `TRUE` when fitting the model. We need this option in order to perform hypothesis tests with the `hypothesis()` function later on.

```{r GLM_propCorrect}

# check default priors
default_prior(prop_correct ~ isAdaptive, data = data_Williams2022_agg)

# change priors for parameters of class b
my_priors_pc <- prior(normal(0,0.25), class = b)

# specify the brmsformula
pc_formula <- bf(
  prop_correct ~ 1 + isAdaptive, 
  sigma ~ 1 + isAdaptive
)

# fit model
fit_pc <- brm(
  formula = pc_formula,
  data = data_Williams2022_agg,
  family = gaussian(),
  cores = 4,
  prior = my_priors_pc,
  sample_prior = TRUE,
  backend = "cmdstanr",
  refresh = 0,
  file = here("models", "model_sdt_glm_pc"),
  file_refit = "never"
)
```

```{r GLM_CowansK}

# check default priors
default_prior(K ~ isAdaptive, data = data_Williams2022_agg)

# change priors for parameters of class b
my_priors_CowansK <- prior(normal(0,1), class = b)

# specify the brmsformula
CowansK_formula <- bf(
  K ~ 1 + isAdaptive, 
  sigma ~ 1 + isAdaptive
)

# fit model
fit_CowansK <- brm(
  formula = CowansK_formula,
  data = data_Williams2022_agg,
  family = gaussian(),
  cores = 4,
  prior = my_priors_CowansK,
  sample_prior = TRUE,
  backend = "cmdstanr",
  refresh = 0,
  file = here("models", "model_sdt_glm_k"),
  file_refit = "never"
)
```

```{r GLM_dprime}

# check default priors
default_prior(dprime ~ isAdaptive, data = data_Williams2022_agg)

# change priors for parameters of class b
my_priors_dprime <- prior(normal(0,1), class = b)

# specify the brmsformula
dprime_formula <- bf(
  dprime ~ 1 + isAdaptive, 
  sigma ~ 1 + isAdaptiv
)

# fit model
fit_dprime <- brm(
  formula = dprime_formula,
  data = data_Williams2022_agg,
  family = gaussian(),
  cores = 4,
  prior = my_priors_dprime,
  sample_prior = TRUE,
  backend = "cmdstanr",
  refresh = 0,
  file = here("models", "model_sdt_glm_dprime"),
  file_refit = "never"
)
```

### 3)  Quantify the strength of evidence for or against an effect of the manipulation.

The `hypothesis` function allows to specify to-be-tested hypotheses for any GLM fitted with `brms`. For this you need to pass the `brmfit` object and a vector of character strings specifying the hypothesis you want to test. Usually you formulate the hypothesis in form of a null hypothesis. In order to use this function you need to set in your model the `sample_prior` option to `TRUE`.

```{r hypothesis_pc}

# proportion correct
hypothesis(fit_pc,
           "isAdaptiveyes = 0")

```

```{r hypothesis_CowansK}

# Cowan's K
hypothesis(fit_CowansK,
           "isAdaptiveyes = 0")

```

```{r hypothesis_dprime}

# dPrime
hypothesis(fit_dprime,
           "isAdaptiveyes = 0")

```

## B - Fit Bayesian-hierarchical SDT model to the data

### 1) Load the **non-aggregated data** of experiment 2 of Williams et al. (2022).

This data (filename: `exercise2_Williams2022_exp2.txt`) contains more detailed data on the number of hits and false alarms (`sum_sayold`) for trials with targets and foils (`was_same_trial`) as test stimuli.

```{r load_data2}
data_Williams2022 <- read.table(here("data","exercise2_Williams2022_exp2.txt"))
head(data_Williams2022)
```

### 2)  Specify the non-linear formula to fit an SDT model with this data.

Use the `brmsformula` function to specify the non-linear formula implementing an SDT model in a GLM. Predict both `dprime` and the `criterion` by the `isAdaptive` manipulation.

```{r sdt_formula}
sdt_formula <- bf(
  sum_saysame | trials(n_trials) ~ -crit + was_same_trial * dprime,
  dprime ~ 1 + isAdaptive + (1 | id),
  crit ~ 1 + isAdaptive + (1 | id),
  nl = TRUE
)
```

### 3)  Fit the SDT model to the data.

Pass the `brmsformula` to the `brm` function and fit a binomial model with a **probit** link function.

Again consider specifying adequate priors for the parameters and plot the model predictions (hint: `conditional_effects()`). Use the `hypothesis` function to test on which parameters of the SDT model the `isAdaptive` manipulation has an effect.

Priors: 

```{r priors_sdt}

# check default priors
default_prior(sdt_formula, data = data_Williams2022, family = binomial(link = "probit"))

# adapt priors
my_priors_sdt <- prior(normal(0,1), class = b, nlpar = crit) +
  prior(normal(0,1), class = b, nlpar = dprime)
```

Model:

```{r fit_sdt}

# fit model
fit_sdt <- brm(
  formula = sdt_formula,
  data = data_Williams2022,
  family = binomial(link = "probit"),
  prior = my_priors_sdt,
  sample_prior = TRUE,
  cores = 4,
  refresh = 0,
  backend = "cmdstanr",
  file = here("models", "model_sdt"),
  file_refit = "never"
)

# plot model predicted effects
conditional_effects(fit_sdt, nlpar = "dprime", effects = "isAdaptive")
conditional_effects(fit_sdt, nlpar = "crit", effects = "isAdaptive")

```

Hypotheses:

```{r}

# test if manipulation had effects on SDT parameters
hypothesis(fit_sdt,
           c("dprime_isAdaptiveyes = 0",
             "crit_isAdaptiveyes = 0"))
```

