---
title: "Analyzing data on the level of cognitive processes - ESCoP 2025"
author: "Gidon T. Frischkorn, Ruhi Bhanap, and Isabel Courage"
output:
  html_document: default
  pdf_document: default
---

```{r chunk_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Exercise 7: Signal Discrimination Model

In this exercise, we will have a look at fitting the Signal Discrimation model to a continous reproduction task varying the set size of to be remembered color patches. 

```{r load_packages, message=F}
# empty environment
rm(list = ls())

# load libraries
library("here")
library("tidyverse")
library("brms")
library("bmm")
```

The data is part of the `bmm` package and can be loaded as `oberauer_lin_2017`.

```{r load_data}
sdm_data <- oberauer_lin_2017
head(sdm_data)
```


Your task is:

1. specify the `sdm` model object using the response error `dev_rad`.

```{r setup_model}
sdm_model <- sdm(resp_error = "dev_rad")
```


2. estimate an `sdm` model evaluating if the parameter `c` or `kappa` changes over set size. Consider entering set size as a numeric variable assuming a linear set size effect

```{r specify_formula}
sdm_data$ss_num <- as.numeric(sdm_data$set_size)

# center on set size 1
sdm_data$ss_num <- sdm_data$ss_num - min(sdm_data$ss_num)

# sort data by predictor for faster fitting
sdm_data_sorted <- sdm_data %>% arrange(ss_num, ID)

sdm_formula <- bmf(
  c ~ 1 + ss_num + (1 + ss_num | ID),
  kappa ~ 1 + ss_num + (1 + ss_num | ID)
)

default_prior(sdm_formula, data = sdm_data, model = sdm_model)
```

3. fit the `sdm` to the data

```{r fit_sdm}
sdm_fit <- bmm(
  formula = sdm_formula,
  data = sdm_data_sorted,
  model = sdm_model,
  # sampler settings
  cores = 4,
  init = 0.5,
  iter = 2000,
  refresh = 100,
  sample_prior = TRUE,
  backend = "cmdstanr",
  file = here("models","exc7_sdm_setsize")
)

brms::pp_check(sdm_fit)
```

4. plot the predicted effects on `c` and `kappa` using the `conditional_effects` function

```{r plot_effects}
conditional_effects(sdm_fit, par = "c", effects = "ss_num")
conditional_effects(sdm_fit, par = "kappa", effects = "ss_num")
```


5. use the `hypothesis` function to test for effects of set size on `c` and `kappa`

```{r test_hypothesis}
ss_effects <- hypothesis(
  sdm_fit,
  c("c_setsize" = "c_ss_num = 0",
    "kappa_setsize" = "kappa_ss_num = 0")
)
ss_effects
plot(ss_effects)
```

If you want compare your results to fitting a two-parameter mixture model on the data. Again, see how the proportion of responses from memory and kappa is affected by the set size variable.

```{r fit_mixture}
mixture_model <- mixture2p(resp_error = "dev_rad")

mixture_formula <- bmf(
  thetat ~ 1 + ss_num + (1 + ss_num | ID),
  kappa ~ 1 + ss_num + (1 + ss_num | ID)
)

mixture_fit <- bmm(
  formula = mixture_formula,
  data = sdm_data,
  model = mixture_model,
   # sampler settings
  cores = 4,
  init = 0.5,
  iter = 2000,
  refresh = 100,
  sample_prior = TRUE,
  backend = "cmdstanr",
  file = here("models","exc7_mixture2p_setsize")
)

pp_check(mixture_fit)

summary(mixture_fit)
```

```{r effects_mixture}
conditional_effects(mixture_fit, par = "thetat", effects = "ss_num")
conditional_effects(mixture_fit, par = "kappa", effects = "ss_num")
```

```{r hypothesis_mixture}
hypothesis(
  mixture_fit,
  c("pMem_ss" = "thetat_ss_num = 0",
    "kappa_ss" = "kappa_ss_num = 0")
)

```


