---
title: "Analyzing data on the level of cognitive processes - ESCoP 2025"
author: "Gidon T. Frischkorn, Ruhi Bhanap, and Isabel Courage"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Solutions: Exercise 5 - EZ-Diffusion Model (Color Judgement taks)

In this exercise, we will have a look at estimating the parameters from the diffusion decision model using the hierarchical Bayesian implementation of the ez-Diffusion Model. This implementation allows for a fast parameter estimation while still providing acceptable recovery of three parameters of interest:

1. drift rate = the rate of evidence accumulation
2. boundary separation = the separation of decision thresholds
3. non-decision time = the time for non-decisional processes.

Your task is to load one of the data sets we prepared for this exercise, either `excercise5_visual_search_task.txt` or `excercise5_color_judgement_task.txt`.

The *Color Judgement* data set contains data of 50 participants that completed a color judgement task. In this task participants saw an array of blue and red dots and should decide if the array contained more blue or red dots. There were two difficulty levels, the easy level in which 60% of the dots were either blue or red, and the difficult level in which 52% of the dots were either blue or red.

The *Visual Search* data set contains data for 33 participants that completed the visual search task. In this task, participants were presented a color cue, a positive (the target color), a negative (the non target color) or neutral (not part of the visual search display). Participants had to look for a "C" that either looked upwards or downwards and report it´s direction with the arrow key.

```{r}
# empty environment
rm(list = ls())

# Load required packages
library(bmm)
library(tidyverse)
library(here)
library(bayestestR)
```

### 1. Prepare data

First, we need to aggregate the data so that we can analyze our data with the `ezdm` using the `bmm` function. For this, we need to calculate the mean reaction time, the variance of the reaction time, the number of correct responses, and the number of trials in each of the four cells of the experimental design: easy - blue correct, easy - red correct, difficult - blue correct, and difficult - red correct.

```{r aggregate data}
ezdm_data <- read.table(here("data","exercise5_color_judgement_task.txt"),
                        header = T, sep = ",")

ezdm_data_agg <- ezdm_data %>% 
  filter(rt > 150, rt < 5000) %>% 
  summarise(
    mrt = mean(rt/1000),
    vrt = var(rt/1000),
    n_correct = sum(correct),
    n_trials = n(),
    .by = c("participant_id", "condition", "correct_response")
  )

ezdm_data_agg$condition <- as.factor(ezdm_data_agg$condition)
contrasts(ezdm_data_agg$condition) <- contr.equalprior
```

### 2. Specify model object

Next, we need to specify the model object to link which data columns are used to identify the `ezdm`.

```{r specify bmmodel}
ezdm_model <- bmm::ezdm(mean_rt = "mrt", 
                        var_rt = "vrt", 
                        n_correct = "n_correct",
                        n_trials = "n_trials")
```

### 3. Specify the BMM formula

Now, we can consider which parameters are affected by the experimental manipulations. First, let's see how the `ezdm` parameters vary as a function of the difficulty manipulation.

```{r specify bmmformula}
ezdm_formula <- bmf(
  drift ~ 1 + condition + (1 + condition | participant_id),
  bound ~  1 + (1 | participant_id),
  ndt ~  1 + condition + (1 + condition | participant_id)
)
```

### 4. Estimate `ezdm` with the `bmm` function

Finally, we can estimate the model using our aggregated data, the `bmmodel` object we created and the `bmmformula` specifying how the model parameters should change over experimental condition.

```{r estimate ezdm}
ezdm_fit <- bmm(
  formula = ezdm_formula,
  data = ezdm_data_agg,
  model = ezdm_model,
  cores = 4,
  sample_prior = TRUE
  #file = here("models","ezdm_colortask")
)
```

We have a look at the model output:

```{r print summary}
summary(ezdm_fit)
```

Moreover, we test whether the estimates for the drift rate and non-decision time are substantial:

```{r}
ezdm_hypothesis <- c(
  driftFX = "drift_condition1 = 0",
  ndtFX = "ndt_condition1 = 0"
)

ezdm_hypothesis_results <- brms::hypothesis(
  ezdm_fit,
  ezdm_hypothesis
)

plot(ezdm_hypothesis_results)
1/abs(ezdm_hypothesis_results$hypothesis$Evid.Ratio)
```
