---
title: "Analyzing data on the level of cognitive processes - ESCoP 2025"
author: "Gidon T. Frischkorn, Ruhi Bhanap, and Isabel Courage"
output:
  html_document: default
  pdf_document: default
---

# Exercise 6: Full Diffusion Decision Model

In this exercise, we will have a look at estimating the parameters from the diffusion decision model using the hierarchical Bayesian implementation of the full Diffusion Decision Model. 

Your task is to load one of the data sets we prepared for this exercise, either `exc4_VisualSearch.txt` or `exc5_ColorJudgement.txt`.

The Color Judgement data set contains data of 50 participants that completed a Color Judgement task. In this task participants saw an array of blue and red dots and should decide if the array contained more blue or red dots. There were two difficulty levels, the easy level in which 60% of the dots were either blue or red, and the difficult level in which 52% of the dots were either blue or red.

In this exercise, we implement the four parameter `ddm` and test if participants were biased in responding with one of the two response options (i.e. blue or red). For this we have to specify a `ddm` with the actual responses - that is blue or red - as the response boundaries. Thus, we will recode the `correct_response` variable into a numeric `response` variable, and then estimate the `drift` separately for the target containing more blue or red dots.

## Solution

```{r setup, message=FALSE}
# Load required packages
library(bmm)
library(tidyverse)
library(here)
```

### 1) prepare the data

First, we need to prepare the data so that we can analyze our data with the `ddm` using the `bmm` function. For this we need to include a numeric  `response` variable coding blue as `0` and red as `1`. Additionally, we need to ensure that our `rt` is stored in seconds. 

```{r aggregate data}
ddm_data <- read.table(here("data","exercise5_color_judgement_task.txt"),
                        header = T, sep = ",")

ddm_data$rt <- ddm_data$rt/1000
ddm_data$resp_num <- ifelse(ddm_data$resp == "red",1,0)
ddm_data$target <- ifelse(ddm_data$correct_response == "red","red","blue")
```

### 2) Specify the `ddm` model object 

Next, we need to specify the model object to link which data columns are used to identify the `ezdm` using the `rt` in seconds and the newly coded `response` variable as the required variables

```{r specify bmmodel}
ddm_model <- ddm(rt = "rt", response = "resp_num")
```


### 3) specify the `bmmformula` with 

Now, we can consider which parameters are affected by the experimental manipulations. Here we assume different `drift` dependent on the target (i.e. are there more red or blue dots) and the difficulty of the task. I also included the condition effect on the non-decision time. Apart from that I only include fixed and random intercepts for the boundary and relative starting point.

```{r specify bmmformula}
ddm_formula <- bmf(
  drift ~ 0 + target:condition + (0 + target:condition || participant_id),
  bound ~ 1 + (1 || participant_id),
  ndt ~ 0 + condition + (0 + condition || participant_id),
  zr ~ 1 + (1 || participant_id)
)

default_prior(ddm_formula, data = ddm_data, model = ddm_model)
```

### 4) estimate the model 

Finally, we use the `bmm` function to fit the full DDM by passing the `ddm` model object and the `bmmformula`. Again, I saved the model results to save time. The `ddm` can take quite some time for parameter estimation. For this data it took about an hour on my machine. 

```{r Estimate ezdm, echo=TRUE, message=FALSE, warning=FALSE}
ddm_fit <- bmm(
  formula = ddm_formula,
  data = ddm_data,
  model = ddm_model,
  cores = 4,
  sample_prior = TRUE,
  backend = "cmdstanr",
  file = here("models","ddm_colortask_exc6")
)
```

### 5) Evalutate results 
 
First, we can have a look at the summary of the fitted `ddm` model object.

```{r print summary}
summary(ddm_fit)
```

Then, we can test some hypothesis:

```{r ddm_hypothesis}
ddm_hypothesis <- c(
  drift_easy = "drift_targetblue:conditioneasy = -drift_targetred:conditioneasy",
  drift_hard = "drift_targetblue:conditionhard = -drift_targetred:conditionhard",
  bias = "zr_Intercept = 0",
  ndt_cond = "ndt_conditioneasy = ndt_conditionhard",
  drift_cond = "(drift_targetblue:conditioneasy - drift_targetred:conditioneasy)/2 =
  (drift_targetblue:conditionhard - drift_targetred:conditionhard)/2"
)

ddm_hypothesis_results <- brms::hypothesis(
  ddm_fit,
  ddm_hypothesis
)

plot(ddm_hypothesis_results)
ddm_hypothesis_results
1/abs(ddm_hypothesis_results$hypothesis$Evid.Ratio)
```

