---
title: "Analyzing data on the level of cognitive processes - ESCoP 2025"
author: "Gidon T. Frischkorn, Ruhi Bhanap, and Isabel Courage"
output:
  html_document: default
  pdf_document: default
---

# Exercise 6: Full Diffusion Decision Model

In this exercise, we will have a look at estimating the parameters from the diffusion decision model using the hierarchical Bayesian implementation of the full Diffusion Decision Model. 

Your task is to load one of the data sets we prepared for this exercise, either `exercise5_VisualSearch.txt` or `exc5_ColorJudgement.txt`.

The Visual Search data set contains data for 33 participants that completed the Visual Search task. In this task, participants were presented a color cue, a positive (the target color), a negative (the non target color) or neutral (not part of the visual search display). Participants had to look for a "C" that either looked upwards or downwards and report it´s direction with the arrow key. 

In this exercise, we implement the four parameter `ddm` and test if participants were biased in responding with one of the two response options (i.e. blue or red). For this we have to specify a `ddm` with the actual responses - that is blue or red - as the response boundaries. Thus, we will recode the `correct_response` variable into a numeric `response` variable, and then estimate the `drift` separately for the target containing more blue or red dots.

## Solution

```{r setup, message=FALSE}
# Load required packages
library(bmm)
library(tidyverse)
library(here)
```

### 1) prepare the data

First, we need to prepare the data so that we can analyze our data with the `ddm` using the `bmm` function. For this we need to include a numeric  `response` variable coding arrowup as `0` and arrowdown as `1`. Convert the reaction times to seconds,

```{r aggregate data}
ddm_data <- read.table(here("data","excercise5_visualsearchtask.txt"),
                        header = T, sep = ",")

ddm_data$rt <- ddm_data$RT/1000
ddm_data$resp_num <- ifelse(ddm_data$response == "arrowdown",1,0)
ddm_data$condition <- factor(ddm_data$condition, 
                             levels = c("negative", "neutral", "positive"))

```

### 2) Specify the `ddm` model object 

Next, we need to specify the model object to link which data columns are used to identify the `ddm` using the `rt` in seconds and the newly coded `response` variable as the required variables

```{r specify bmmodel}
ddm_model <- ddm(rt = "rt", response = "resp_num", version = "four_par")
```


### 3) specify the `bmmformula` with 

Now, we can consider which parameters are affected by the experimental manipulations. Here we assume different `drift` and `ndt` dependent on the the type of cue (positive/negative/neutral) and a `drift` based on the correct response. Apart from that I only include fixed and random intercepts for the boundary and relative starting point.

```{r specify bmmformula}
ddm_formula <- bmf(
  ddm_formula <- bmf(
  drift ~ 0 + correct:condition + (0 + correct:condition || participant),
  bound ~ 1 + (1 || participant),
  ndt ~ 0 + condition + (0 + condition || participant),
  zr ~ 1 + (1 || participant)
)

default_prior(ddm_formula, data = ddm_data, model = ddm_model)
```

### 4) estimate the model 

Finally, we use the `bmm` function to fit the full DDM by passing the `ddm` model object and the `bmmformula`. Again, I saved the model results to save time. The `ddm` can take quite some time for parameter estimation. For this data it took about an hour on my machine. 

```{r Estimate ezdm, echo=TRUE, message=FALSE, warning=FALSE}
ddm_fit <- bmm(
  formula = ddm_formula,
  data = ddm_data,
  model = ddm_model,
  cores = 4,
  iter = 2000,
  sample_prior = TRUE,
  backend = "cmdstanr",
  file = here("models","ddm_colortask_exc6")
)
```

### 5) Evalutate results 
 
First, we can have a look at the summary of the fitted `ddm` model object.

```{r print summary}
summary(ddm_fit)
```

Then, we can test some hypothesis:

```{r ddm_hypothesis}
ddm_hypothesis <- c(
  # Drift rates: comparing arrow down vs arrow up within each condition
  drift_neg = "drift_correctarrowdown:conditionnegative = -drift_correctarrowup:conditionnegative",
  drift_neu = "drift_correctarrowdown:conditionneutral = -drift_correctarrowup:conditionneutral",
  drift_pos = "drift_correctarrowdown:conditionpositive = -drift_correctarrowup:conditionpositive",
  
  # Starting point / bias
  bias = "zr_Intercept = 0",
  
  # Non-decision time comparisons
  ndt_neg_neu = "ndt_conditionnegative = ndt_conditionneutral",
  ndt_neu_pos = "ndt_conditionneutral = ndt_conditionpositive"
)


ddm_hypothesis_results <- brms::hypothesis(
  ddm_fit,
  ddm_hypothesis
)

plot(ddm_hypothesis_results)
ddm_hypothesis_results
1/abs(ddm_hypothesis_results$hypothesis$Evid.Ratio)

ddm_hypothesis_results <- brms::hypothesis(
  ddm_fit,
  ddm_hypothesis
)

plot(ddm_hypothesis_results)
ddm_hypothesis_results
1/abs(ddm_hypothesis_results$hypothesis$Evid.Ratio)
```
